Reproducible development environments with rix

### 2024-01-29 paper.qmd
fixes #2
with_nix

We also provide a way to start a temporary shell with [R]{.proglang} and
[rix]{.proglang} available, which makes it possible to start building
reproducible environments on a system that does provide [Nix]{.proglang} but not
[R]{.proglang}.

@@

@@

```{r}
library(rix)

>>>>

## Additional features of the rix package

@@

The final feature that we provide is the ability to run one arbitrary function
from what we call a *subshell*. Suppose that you are running the latest version
of R, and that you wish or need to run one particular function from an older
release of R, or from an older package. Using the `with_nix()` function, it is
possible to run this single function inside the right envirnoment, and recover
its output in the main session. To illustrate, consider the following two lines
of code:

```{r, eval = FALSE}
set.seed(1234)
sample(seq(1, 10), 5)
```

On versions of R released after version 3.6, these two lines return the
following vector, `r set.seed(1234);sample(seq(1, 10), 5);set.seed(NULL)`, in
versions prior to 3.6, the result will be different. This is a documented change
as shown in the R version 3.6 release
notes.^[https://stat.ethz.ch/pipermail/r-announce/2019/000641.html]

When trying to reproduce a study made on versions prior to 3.6, a replicator
will need to take this into account. Using the `with_nix()` function,
the original results can be replicated even from the latest version of R.

The first step is to create an environment with the old release of R:

```{r}
path_subshell <- tempdir()

rix(
  r_ver = "3.5.3",
  overwrite = TRUE,
  project_path = path_subshell
)
```

We recommend `path_subshell` to be a subfolder of the project. Using `rix()`
we generate a `default.nix` file that will provide a shell with R version 3.5.3
when built. We then use `rix_init()`, which generates an `.Rprofile` file in
the same subfolder as where the subshell's `default.nix` is located. This `.Rprofile`
file ensures that the R session that will be spawned from this subshell does not
get any interference from the main session it gets called from:

```{r}
rix_init(
   project_path = path_subshell,
   rprofile_action = "overwrite",
   message_type = "simple"
 )
```

We can now use this shell to generate our vector of random numbers using `with_nix()`:

```{r}
out_subshell <- with_nix(
    expr = function(){
      set.seed(1234)
      a <- sample(seq(1, 10), 5)
      set.seed(NULL)
      return(a)
    },
    program = "R",
    exec_mode = "non-blocking",
    project_path = path_subshell,
    message_type = "simple"
  )

```

The `expr` argument of `with_nix()` can be arbitrarly complex, but must
be wrapped inside a `function()` call. The value of `out_subshell` is
`r out_subshell`, as is indeed different than the one generated from a
recent version of R.

@@

```{r, echo = FALSE}
library(rix)
